---
output: github_document
---

## population.R

The `population.R` script generates an artificial population for the epidemiological
model. Here we use the data from Toth *et al.* (2021), in particular, the distribution
of household sizes in Salt Lake City. The simulation has two parts: sampling
the households and interconnecting individuals between the households.

### Simulating households

1. Set a desired population size, `N_desired`, number of locations, `N_locations`,
and number of visits each individual does, `N_avg_locations`.

2. Using the data from Toth *et al.*, retrieve a weighed sample of households
using the observed frequencies as weights.

3. Select the first `H` households sampled such that the sum of their sizes is
at most `N_desired`. 

4. If the sum of the household sizes is less than `N_desired`, add assign the 
difference to the last household.

### Interconnecting households

There are many strategies possible to interconnect individuals between households.
At this time, we will connect individuals based on co-afiliation networks, in particular,
we will assign individuals to locations at random and create ties between those
who co-occurr.

One problem with this approach is that, as the population size increases, the
average number of connections grows as well. Having a network with a significantly
high mean degree is not desired as in reality density decreases with size.

To formally analyze this, we can calculate the expected mean degree as a function
of the number of locations in the system and mean number of locations individuals
go to. Let $i \in N$ denote a population of size $n$, $j \in M$ the set of $m$ 
locations, and $\mu_v$ the mean number of visits per individual, with $v\sim\mbox{Binom}(m, \mu_v/m)$. Assuming all. Let $p=\mu_v/m$; the probability that individuals $i$ and $k$ visit the same place
can be computed as follows:

$$
\begin{align*}
P(i\leftrightarrow k) & = 1 - P(\neg i \leftrightarrow k) \\
& = 1 - \prod_{j\in M} P(\mbox{not both in }j) \\
& = 1 - \prod_{j\in M} [1 - P(\mbox{both in }j)] \\
& = 1 - \prod_{j\in M} [1 - P(i\in j, k\in j)] \\
& = 1 - \prod_{j\in M} [1 - P(i\in j) P(k\in j)] \\
& \mbox{Since both are symmetrical} \\
& = 1 - \prod_{j\in M} (1 - (\mu_v/m)^2) \\
& = 1 - [1 - (\mu_v/m)^2]^{m}
\end{align*}
$$

This way, the expected degree of the system is then

$$
\begin{align*}
\mathbb{E}(degree) & = \left[1 - [1 - (\mu_v/m)^2]^{m}\right]\times(N - 1)
\end{align*}
$$
Furthermore, the degree is distributed binomial with mean $1 - [1 - (\mu_v/m)^2]^m$
and size $N - 1$ (which explains the result we obtained with the expected degree).

As the number of individuals grows so does mean degree. 
In this case, if we wire the individuals through a bipartite graph, then we have
to take into consideration that individuals may tend then to have a large number
of connections in large systems, which may not reflect reality.

```{r figure-levelplot, echo = FALSE, fig.cap="Expected degree as a function of Number of locations\nand average visits"}
library(ggplot2)

# Expected degree
expected_deg <- function(n, m, mu_v) {
  (n - 1) * (1 - (1 - (mu_v/m)^2) ^ m)
} 

N <- c(1000, 10000, 50000, 100000)
nout <- 50
Mu_v <- seq(from = .1, to = 5, length.out = nout)
dat <- NULL
for (n in N) {
  M <- rev(floor((seq(from = 100, to = 500, length.out = nout))))
  D <- outer(Mu_v, M, \(mu_v, m) expected_deg(n, m, mu_v))
  
  dat <- rbind(dat, data.frame(
    n      = n,
    degree = as.vector(D),
    m      = sort(rep(M, nout)),
    mu_v   = Mu_v
  ))
}

labs_n <- labeller(n = setNames(
  paste(
    formatC(
      N, big.mark = ",", format="g", digits = 10,
      width = nchar(sprintf("%i", max(N))) + 2
      ), " agents"
    )|> trimws(which = "both"), N))

ggplot(dat, aes(x = m, y = mu_v, z = degree)) +
    facet_wrap(~n, labeller = labs_n) + 
    geom_contour_filled(
      breaks = round(exp(seq(log(1), log(10000), length.out = 15)))
    ) +
  labs(
    x = "Number of Locations",
    y = "Average Number of Visits\n(binomial dist)",
    fill = "Mean Degree"
    )
```

From the figure we can see that, for example, in a system with 50,000 agents,
400 locations, and each visiting an average of two locations, the expected degree
is `r prettyNum(round(expected_deg(50000, 400, 2)), big.mark = ",")`.

## Appendix

### Checking predicted mean degree

The following code-block performs a brief simulation study to validate this result:

```{r simul-example, cache = TRUE}
library(data.table)
n    <- 400
m    <- 50
mu_v <- 1

expected_deg(n, m, mu_v)

# Running a simulation
set.seed(7123)
N     <- 1:n
nsims <- 10000
M     <- 1:m

Visits <- array(runif(n * m * nsims) < mu_v/m, dim = c(n, m, nsims))

sim <- parallel::mclapply(1:nsims, \(i) {
  
  # Sampling per individual
  visits <- which(Visits[,,i], arr.ind = TRUE)

  visits <- data.table(
    id = visits[,1],
    loc = visits[,2]
  )

  # Creating the bipartite graph  
  visits <- merge(
    visits[, .(ego = id, loc)],
    visits[, .(alter = id, loc)],
    by = "loc", allow.cartesian = TRUE, all = TRUE
  )[, loc:=NULL] |> unique()
  
  # Retrieving the edgelist
  visits <- visits[, .(ego = fifelse(ego > alter, ego, alter), alter = fifelse(ego > alter, alter, ego))]
  visits <- unique(visits)[ego != alter,]
  
  # Calculating the degree
  sum(table(visits[,c(ego, alter)]))/n
  
}, mc.cores = 6L) |> unlist()

hist(sim)
mean(sim)
```

