---
output: github_document
---

# Model 0 for Salt Lake City

This simulation model features the following:

1. Five different states: Susceptible, Exposed, Infected, Recovered, and Removed.

2. The population size is 10,000 individuals, all connected at the household level,
 and between households.

3. A baseline probability of infection is .9.

4. Daily recovery rate for an infected agent is `1/7`.

5. Daily death rate for an infected agent is .01.

6. There is one virus that starts with 5% of the population exposed.

7. The simulation lasts 100 days, and replicated a thousand times.

The implementation details can be found in [00main.cpp](00main.cpp). After
compilation, the execution of the program results in the following output:

```{bash}
./00main.o
```

Statuses' trajectories have the following distribution:

```{r posterior-dist, dev="svg", echo = FALSE}
library(data.table)
library(ggplot2)
dat <- lapply(list.files("results",  pattern = "total_hist", full.names = TRUE), fread)
dat <- Map(\(d, i) d[, r := i], d = dat, i = seq_along(dat) - 1) |>
    rbindlist()

summ <- dat[, .(
    `2.5%`  = quantile(counts, probs = .025),
    Mean    = mean(counts),
    `97.5%` = quantile(counts, prob = .975)
    ), by = .(date, status)][
        date %inrange% c(1, 100)][
            status != "Susceptible"]

setorder(summ, status, date)

ggplot(summ, aes(x = date, colour = status)) +
    geom_ribbon(aes(
        y = Mean, ymin = `2.5%`, ymax = `97.5%`,
        colour = status
        ), alpha = .3) +
    geom_line(aes(y = Mean)) +
    labs(
        x = "Day", y = "Counts",
        title = "Distribution of cases after 1,000 replicates"
        ) 
```

Distribution of the mean reproductive number

```{r rep-num, dev="svg", echo = FALSE}
repnum <- lapply(
    list.files("results", pattern = "repro", full.names = TRUE), fread
)

repnum <- sapply(repnum, \(x) x[, mean(rt)]) |>
    data.table() |> setnames("rt")
ggplot(repnum, aes(x = rt)) +
    geom_histogram() + 
    labs(x = "Reproductive Number", y = "Count",
    title = "Distribution of Reproductive number after 1,000 replicates")
```

Distribution of the Rt

```{r rt-num, dev="svg", echo = FALSE}
transm <- lapply(
    list.files("results", pattern = "transm", full.names = TRUE), fread
)

transm <- lapply(transm, \(x) {
    x[, .(rt = .N), by = .(variant, source_exposure_date, source)]
})

transm <- lapply(transm, \(x) {
    x[, .(rt = mean(rt)), by = .(variant, source_exposure_date)]
}) 

transm <- Map(\(d, i) d[, r := i], d = transm, i = seq_along(transm) - 1) |>
    rbindlist()

setnames(transm, "source_exposure_date", "date")

summ_transm <- transm[, .(
    `2.5%`  = quantile(rt, probs = .025),
    Mean    = mean(rt),
    `97.5%` = quantile(rt, prob = .975)
    ), by = .(date)][date %inrange% c(1, 100)]

setorder(summ_transm, status, date)

ggplot(summ_transm, aes(x = date)) +
    geom_ribbon(aes(
        y = Mean, ymin = `2.5%`, ymax = `97.5%`
        ), alpha = .3) +
    geom_line(aes(y = Mean)) +
    labs(
        x = "Day", y = "Rt",
        title = "Distribution of Rt after 1,000 replicates"
        )
```
