---
format: gfm
title: Example of transmission network
author: George G. Vega Yon
date: 2024-07-25
---

We usee a network SIR model with the following parameters:

- Population size (`N`): 100.
- Prevalence: 1/`N`.
- Transmission rate: 0.1.
- Recovery rate: 1/7.

The underlying network is a homophilic network on age with average degree 11.12.

```{r}
#| label: network
library(network)
library(ergm) # Simulating an ERGM network
library(netplot)

N <- 100
set.seed(121)
age <- sample.int(20, N, replace = TRUE) + 15

g   <- network.initialize(N, directed = FALSE)
g %v% "age" <- age
g <- simulate_formula(g ~ edges + absdiff("age", pow=1.5), coef = c(0, -1))

nplot(g, vertex.color = ~ age)
```

```{r}
el <- network::as.edgelist(g)

library(epiworldR)

# Building the model
N <- 100
mymodel <- ModelSIR(
  name = "COVID-19",
  prevalence = 2/N,
  transmission_rate = 0.1,
  # incubation_days = 2,
  recovery_rate = 1/7
)

agents_from_edgelist(
  model = mymodel,
  source = as.integer(el[,1] - 1L),
  target = as.integer(el[,2] - 1L),
  size = N,
  directed = FALSE
)
```

We now run and visualize the model output

```{r}
#| label: run-the-model
run(mymodel, ndays = 50)
summary(mymodel)
plot_incidence(mymodel)
```

The network visualization is done using `netplot`. We use the same layout for both the transmission and contact networks.

```{r}
library(igraph)
tnetwork <- get_transmissions(mymodel)

tnet <- graph_from_data_frame(
  d = tnetwork[, c("source", "target")],
  vertices = data.frame(name = (1:N) - 1)
)

set.seed(44)
lo <- layout_nicely(tnet)

p1 <- nplot(g, layout = lo, skip.vertex = TRUE)
p2 <- nplot(tnet, layout = lo, skip.vertex=TRUE)

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

Finally, we save the networks using rds files.

```{r}
#| label: save-networks
saveRDS(g, "contact-network.rds")
saveRDS(tnet, "transmission-network.rds")
```